
---
title: "Predicting Box Office Bombs: A Logistic Regression Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
format:
  pdf:
    toc: true
    number-sections: true
bibliography: references.bib
---

# Introduction

This report explores the factors influencing whether movies become "box office bombs" using logistic regression. We utilize a dataset containing various attributes such as runtime, critic scores, budget, and distribution types. Predicting movie success or failure aids studios in decision-making to mitigate financial risks [@james2021, @kuhn2022].

## Research Question

What factors significantly predict whether a movie will fail financially (become a box office bomb)?

## Dataset Overview

```{r}
#| echo: true
source("helpers.R")

df_movie <- read_and_prepare_movie_data("movie_select.csv")

df_movie %>% 
  select(title, budget, domestic, runtime, audience_score, critic_score, is_bo_bomb) %>% 
  arrange(-budget) %>% head(7)
```

# Exploratory Data Analysis (EDA)

## Distributions and Insights

We first examine distributions of key variables, helping us understand the nature of our predictors.

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 7
#| fig-align: center
library(ggplot2)
library(patchwork)

distribution_plots(df_movie)
```

From the plots, we notice distinct patterns in runtime, critic scores, indie movies, and box office bombs.

## Relationships and Correlations

Investigating correlations allows us to identify multicollinearity or redundant variables.

```{r}
#| echo: true
correlation_matrix(df_movie)
```

Runtime and critic scores appear moderately correlated with box office performance.

# Methodology

## Logistic Regression Model

To predict box office failures, we use logistic regression, which models binary outcomes (bomb or not).

The logistic regression equation is defined as follows [@james2021]:

$$
\log\left(\frac{p}{1 - p}\right) = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \dots + \beta_k X_k
$$

where $p$ is the probability of a movie becoming a box office bomb.

## Model Specification

We include predictors based on their relevance and data availability:

- Runtime
- IMDb average rating
- Audience score
- Critic score
- Independent distributor status (Indie)
- Genre indicators (Action, Animation)

# Results

## Model Estimation

We fit the logistic regression model and interpret the results:

```{r}
#| echo: true
logit_model <- glm(is_bo_bomb ~ runtime + imdb_avg_rating + 
                     audience_score + critic_score + is_indie + 
                     ge_action + ge_animation,
                   data = df_movie,
                   family = binomial)

summary(logit_model)
```

The output above highlights significant predictors influencing box office performance.

## Model Evaluation

To evaluate our logistic model, we use McFadden's R² as a goodness-of-fit measure:

```{r}
#| echo: true
library(pscl)
pscl::pR2(logit_model)
```

McFadden's R² above 0.2 generally indicates good fit [@hosmer2013].

## Interpretation of Key Predictors

- **Critic Score:** Higher critic scores significantly reduce the likelihood of a bomb.
- **Runtime:** Longer runtimes modestly increase failure risks.
- **Independent Distribution:** Independent movies show increased risks, possibly due to lower marketing budgets.

# Discussion

The results suggest that critical reception and runtime play vital roles. Studios might reconsider overly lengthy films or films without mainstream backing if minimizing financial risk is essential.

# Conclusion

Our logistic regression analysis effectively identified crucial variables that predict box office failures, thus providing valuable insights for film industry stakeholders.

# References
